Adaptive Video Streaming with Optimized Bitstream Extraction and PID-based Quality Control

### 2015-12-02 tmm.tex
Edit the part of PID parameter selection

As with other PID control systems in the industry, the selection of the three parameters is implementation-dependent and most of the time manual tuning with trial and error is required to achieve optimal results. Knowing the impacts of each parameter is useful in the practice of parameter selection. Generally,  the proportional part in PID controller affects the system's stability. If the proportional gain $K_p$ is too high, the system can become unstable. However, a low $K_p$ would result in long rising time that makes the system response slowly. In addition to eliminating steady-state error, the integral term also accelerates the movement of the process towards setpoint, which means increasing $K_i$ can reduce rising time. However, since the integral part accumulates errors from the past, it can cause the present value to overshoot the setpoint value, and too much $K_i$ will lead to instability as well. This can be compensated by the derivative part, which predicts system behaviour and has the effect of reducing the overshoot, increasing the stability, and improving the settling time (time it takes to reach setpoint after a disturbance). The big problem with derivative part is that it is very sensitive to noise, therefore industrial practices indicates that derivative gain $K_d$ should always stay low.

For the proposed quality control algorithm, Ziegler--Nichols method \cite{Ziegler42} is used to select the parameters to obtain better performance. This method is performed by setting integral and derivative parameters to zero and then tuning $K_p$ to a value $K_u$ until output begins to oscillate with constant amplitude. Assuming the oscillation period is $T_u$, we set $K_p = 0.6K_u$, $K_i = 2K_u/T_u$, and $K_d = K_u \cdot T_u/8$. $K_p$, $K_i$, and $K_d$ are then normalized to get the final parameters. In the following experiments, parameters are chosen as $K_p = 0.22$, $K_i = 0.73$, $K_d = 0.05$ by this method.

In this work we simply use the standard Ziegler--Nichols method which already takes above impacts into consideration and produces acceptably good results. Adopting more sophisticated or automated techniques to further optimize the parameters will be done in the future. It should also be noted that, in the video quality control algorithm only the bitrate and the bandwidth are relevant, not the video content. Therefore once the parameters are chosen they should apply well to different video streams.

>>>>

As with other PID control systems in the industry, the selection of the three parameters is implementation-dependent and most of the time manual tuning with trial and error is required to achieve optimal results. Each parameter has impacts on different aspects of the system, and knowing them is useful in the practice of parameter selection. Generally,  the proportional part in PID controller affects the system's stability. If the proportional gain $K_p$ is too high, the system can become unstable. However, a low $K_p$ would result in long rising time that makes the system response slowly. In addition to eliminating steady-state error, the integral term also accelerates the movement of the process towards setpoint, which means increasing $K_i$ can reduce the rising time. However, since the integral part accumulates errors from the past, it can cause the present value to overshoot the setpoint value, and too much $K_i$ will lead to instability as well. These can all be compensated by the derivative part, which predicts system behaviour and has the effect of reducing the overshoot, increasing the stability, and improving the settling time (time it takes to reach setpoint after a disturbance). The big problem with derivative part is that it is very sensitive to noise, therefore industrial practices indicates that derivative gain $K_d$ should usually stay low.

For the proposed quality control algorithm, Ziegler--Nichols method \cite{Ziegler42} is used to select the parameters for better performance. This method is performed by setting integral and derivative parameters to zero and then tuning $K_p$ to a value $K_u$ until output begins to oscillate with constant amplitude. Assuming the oscillation period is $T_u$, we set $K_p = 0.6K_u$, $K_i = 2K_u/T_u$, and $K_d = K_u \cdot T_u/8$. $K_p$, $K_i$, and $K_d$ are then normalized to get the final parameters. In the following experiments, parameters are chosen as $K_p = 0.22$, $K_i = 0.73$, $K_d = 0.05$ by this method. With these parameters, the PID-controlled streaming system is stable and quickly begins to follow the bandwidth change, as shown by the experimental results in next section.

We choose to use the standard Ziegler--Nichols method because it already takes each parameter's impacts into consideration and produces acceptably good results. Adopting more sophisticated or automated techniques to further optimize the parameter selection and controller performance will be our future work. Note that, in the video quality control algorithm only the bitrate and the bandwidth are relevant, not the video content. Therefore once the parameters are chosen they should apply well to different video streams.
